---
title: "Web Scraping"
author: Peter Claussen
date: 2/11/2020
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# 1 A Gentle Introduction

## The Problem

- Visit the NIST Statistical Reference Data Set [](https://www.itl.nist.gov/div898/strd/anova/SiRstv.html)
- Use the data in [Data File in Two-Column Format](https://www.itl.nist.gov/div898/strd/anova/SiRstv.dat)
- Reproduce the [image](https://www.itl.nist.gov/div898/strd/anova/SiRstv.gif)


## Manual Solution

- Download file to local drive
- Edit in text editor to remove comments
- Read using standard data table functions

# R Solution

```{r,fig.width=8,fig.height=7}
tbl <- read.table('./SiRstv.dat',header=TRUE)
par(mar=c(5,6,2,2))
plot(Resistance ~ Instrument, data=tbl, 
     xlab='Instrument', ylab='Resistivity, ohm*cm',pch=16)
```


# Next Steps

We're a long way from web scraping. We had to download a file from the web to edit it into an appropriate format, deleting documentation. That's acceptable for a small exmaple, more interesting data will be larger and dynamic.

Our next step towards web scraping will be to read files directly from the web. 


## Skipping lines

Since we know we deleted 60 lines of text from the original, we can simply skip lines:

```{r,fig.width=8,fig.height=7}
tbl <- read.table('https://www.itl.nist.gov/div898/strd/anova/SiRstv.dat',skip=60)
names(tbl) <- c('Instrument','Resistance')
plot(Resistance ~ Instrument, data=tbl, 
     xlab='Instrument', ylab='Resistivity, ohm*cm',pch=16)
```


Alternatively,

```{r,fig.width=8,fig.height=7}
tbl <- read.table('https://www.itl.nist.gov/div898/strd/anova/SiRstv.dat',
                  header=TRUE,skip=59,fill=TRUE)
names(tbl) <- c(names(tbl)[2:3],'Blank')
plot(as.formula(paste(names(tbl)[2],'~',names(tbl)[1])), data=tbl,
     xlab='Instrument', ylab='Resistivity, ohm*cm',pch=16)
```

# Learning the structure of a file

This method still requires we know something about the contents of the file. How can we discover contents of the file, without loading the entre file?

## One approach

```{r}
lines <- readLines('https://www.itl.nist.gov/div898/strd/anova/SiRstv.dat')
for(i in 1:length(lines)) {
  if(nchar(lines[i])>0) {
    current.row <- read.table(text=lines[i])
    if(all(apply(current.row,2,is.numeric))) {
      break
    }
  }
}
i
```

So,

```{r,fig.width=8,fig.height=7}
tbl <- read.table('https://www.itl.nist.gov/div898/strd/anova/SiRstv.dat',skip=i-1)
names(tbl) <- c('Instrument','Resistance')
par(mar=c(5,6,2,2))
plot(Resistance ~ Instrument, data=tbl, 
     xlab='Instrument', ylab='Resistivity, ohm*cm',pch=16)
```
...

... or

```{r,fig.width=8,fig.height=7}
tbl <- read.table(text=paste(lines[(i+1):length(lines)],collapse='\n'))
names(tbl) <- c('Instrument','Resistance')
par(mar=c(5,6,2,2))
plot(Resistance ~ Instrument, data=tbl, 
     xlab='Instrument', ylab='Resistivity, ohm*cm',pch=16)
```



Exercises
========================================================

# 1

Read a data table from [Data File in Table Format](https://www.itl.nist.gov/div898/strd/anova/SiRstvt.dat)

#2

Repeat with a different data sets from https://www.itl.nist.gov/div898/strd/general/dataarchive.html



# Python in RStudio


(set `eval=TRUE` to run the following chunks)

In order to run `python` from RStudio, you might need to match versions, in the command line
```
$ python -V
$ which python
```

and then choose specific option if necessary

```{r,eval=FALSE}
library(reticulate)
use_python('/opt/anaconda3/bin/python')
py_config()
```

Then, python output should be redirected to R Markdown

```{python,eval=FALSE}
import pandas
tbl = pandas.read_table('./SiRstv.dat',sep='\s+')
tbl.plot.scatter(x='Instrument', y='Resistance')
```

```{python,eval=FALSE}
import matplotlib.pyplot as plt
plt.plot(tbl['Instrument'], tbl['Resistance'],'ko')
plt.ylabel('Resistivity, ohm*cm')
plt.xlabel('Instrument')
plt.show()
```
